---
title: "rankclust: Fit a Bayesian, Rank-Clustered Bradley-Terry-Luce Model to Ordinal Comparison Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  rmarkdown.html_vignette.check_title = FALSE
)
```

This package implements the Rank-Clustered Bradley-Terry-Luce model proposed by [Pearce and Erosheva (2025)](https://www.cambridge.org/core/journals/psychometrika/article/bayesian-rankclustering/07B890BEA1E9F4C24A5751A1C64A5A87). This Bayesian model for ordinal comparison data of many different types (such as complete or partial rankings, pairwise comparisons, etc.) infers an overall ranking of objects while permitting some objects to be "rank-clustered", i.e., assigned equal rank in the overall ranking.

For more details on how to use this package, see the [Tutorial](https://pearce790.github.io/rankclust/articles/tutorial.html) and [Reference](https://pearce790.github.io/rankclust/reference/index.html) pages on this website. 

The package is currently in development. You may install the package from Github using the following code:
```{r, eval=FALSE}
## Development (Github) version
# install.packages("devtools") # uncomment if you haven't installed 'devtools' before
devtools::install_github("pearce790/rankclust")
```

After installation, load the package with the following code:
```{r}
library(rankclust)
```

# Funding

This project was supported by the National Science Foundation under Grant No. 2019901.
