---
title: "rankclust: Fit a Bayesian, Rank-Clustered Bradley-Terry-Luce Model to Ordinal Comparison Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  rmarkdown.html_vignette.check_title = FALSE
)
```

This package implements the Rank-Clustered Bradley-Terry-Luce model proposed by Pearce and Erosheva (2024+). More specifically, it fits a Bayesian model for ordinal comparison data of many different types (such as complete or partial rankings, pairwise comparisons, etc.) such that some objects being ranked may be "rank-clustered", i.e., assigned equal rank in the overall ranking.

For more details on how to use this package, see the [tutorial](https://pearce790.github.io/rankclust/articles/tutorial.html). 

The package is currently in beta testing. You may install the package from Github using the following code:
```{r, eval=FALSE}
## Beta (Github) version
# install.packages("devtools") # uncomment if you haven't installed 'devtools' before
devtools::install_github("pearce790/rankclust")
```

After installation, load the package with the following code:
```{r echo=FALSE,results='hide',message=FALSE,warning=FALSE}
devtools::install_github("pearce790/rankrate")
```
```{r}
library(rankclust)
```

# Funding

This project was supported by the National Science Foundation under Grant No. 2019901.
